A Motion Capture and Imitation Learning-based Approach to Robot Control

### 2022-06-29 tex-arxiv/article.tex
fix vector notation to be consistent - first pass

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amsmath}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{cleveref}       % smart cross-referencing
\usepackage{lipsum}         % Can be removed after putting your text content
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{doi}
\usepackage{subcaption}

@@

In this article, an \emph{agent} can be taken to mean the part of the system that acts based on state or observation information $s$ in an \emph{environment}, according to a \emph{policy} $\pi$ which is specified by a parametric \emph{model} -- a function $\pi_{\theta}(s)$ with parameters $\theta$ tuned in optimization and produces an output \emph{action} $a$. For all practical intents and purposes, this means that references to the agent, model, and policy are almost interchangeable in most contexts. A formalism common in both reinforcement and imitation learning contexts is the \emph{Markov decision process} (MDP), formally given by \citep{attia2018global}

\begin{equation}
	MDP = (S,A,T,R,I)
\end{equation}
where $S$ is the set of states $s$, $A$ is the (formally discrete) set of actions $a$, $T:S \times A \rightarrow S$ is a state transition function that encapsulates the environment, $R:S \rightarrow \mathbb{R}$ is a reward function associated with each state and $I = p\left(s_0 \in S\right)$ represents the initial state distribution. In many imitation learning-related cases a reward function need not be specified or considered. Moreover, if one is willing to break with the strict formal definition and give up the use of mathematical tools defined only on finite probability distributions, continuous action spaces can also be considered.

@@

\begin{equation}
	t, k, n \in \mathbb{N}; \pi_{\theta}:\lbrace {(s_n)_{t-k}^t} \rbrace \rightarrow A
\end{equation}
which, by allowing for input sequences of variable length, may become a function defined on the entire known state history:

\begin{equation}
	\pi_{\theta}:\lbrace {(s_n)_1^t} \rbrace \rightarrow A
\end{equation}

@@

\begin{equation}
	\pi_{\theta}:\lbrace {(s_n)_1^t} \rbrace \rightarrow S
\end{equation}
which, when running the model on its prior output, is equivalent to sequence-building tasks encountered in domains such as text generation.

@@

After each recording session, an entirely unstructured corpus is available. It contains timestamped pose observations collected at whatever intervals they happened to have been generated by the tracking software. Each observation only contains information about a single tracked body. A peculiarity of this particular recording method is that observations about each body tend to arrive in bursts:
\begin{equation}
    s^a_{t1}, s^a_{t2}, ..., s^b_{tm-1}, s^b_{tm}
\end{equation}
where $s^a_{t \in \mathbb{R}}$ corresponds to the observed state of object $a$ at time $t$.
Thus, to obtain sequential data in discrete time, it is necessary to resample the positions and orientations of all relevant bodies at a constant time step:
\begin{equation}
    (s^a_{1}, s^b_{1}, s^c_{1}), ..., (s^a_{k}, s^b_{k}, s^c_{k}) = s_1, ..., s_k
\end{equation}

@@

\begin{equation}
	(s_{t_{start}}, ... , s_{t_{start}+steps}), steps \in \mathbb{N}
\end{equation}

@@

\begin{equation}
	(\mathbf{r}_{t+1}, \mathbf{q}_{t+1}, g_{t+1}) = \pi_{\theta}\left(\frac{t}{f_{sample}}, \mathbf{r}_{t}, \mathbf{q}_{t}, g_{t}, \mathbf{r}^{target}_t \right)
\end{equation}
where $\mathbf{r}_{t}, \mathbf{q}_{t}, g_{t}$ represent the end effector translation vector, orientation quaternion and gripper actuator signal respectively at time step $t$ in both the input and output. The input is augmented with a time/phase signal (discrete time step divided by the sampling frequency, in this case $100Hz$) relative to the start of the demonstration or generated trajectory, as well as the target coordinate vector $\mathbf{r}^{target}_t$ which corresponds to the extrapolated target coordinates in the demonstration data set and commanded throw coordinates at inference. The time signal was added to the input as it was found that trajectories generated by feedforward networks were liable to diverge without it, and the data sets thus modified were used for all training thereafter. The gripper control signal has values $\lbrace 0,1 \rbrace$ in the training data set.

@@

\begin{equation}
	\mathcal{D}_{eval} = \left ( (\tau^d_{1}, \tau^g_{1}), ..., (\tau^d_{k}, \tau^g_{k}) \right )
\end{equation}
\begin{equation}
	\tau^d, \tau^g = (s^d_1, ..., s^d_m), (s^g_1, ..., s^g_m); s^d_1 = s^g_1 
\end{equation}
where $\mathcal{D}_{eval}$ refers to the evaluation data set of a single model with respect to either the validation or test demonstration set, but $\tau^d_i, \tau^g_i$ are the demonstration and generated trajectories (sequences of state observations $s_j$) sharing the same initial state respectively.

@@

\begin{equation}
	f_{stepwise}(\mathcal{D}_{eval}) = f: \left [ \sum_{\tau^d, \tau^g \in \mathcal{D}_{eval}}\sum_{i=1}^m \left ( g:s^d_i \times s^g_i \rightarrow \mathbb{R} \right ) \right ] \rightarrow \mathbb{R}
\end{equation}
where the inner function $g$ corresponds to metrics such as position error, rotation error (quaternion angular distance), or categorical cross-entropy in the release signal. Finally, a release error metric was computed trajectory-wise. To do this, the release point of each trajectory was found, the corresponding position found and the velocity vector estimated. When considering ways to combine these two terms into a single quantitative error estimator, it was decided that a simple ballistic extrapolation and resulting miss distance along the ground plane would serve as a decent first-order approximation of throw accuracy \emph{vis-à-vis} the training or validation data set.

@@

\begin{equation}
	r_{intersect} = \left ( x(t_{intersect}), y(t_{intersect}), z(t_{intersect}) \right )
\end{equation}
and the trajectory-wise throw error metric computed by:

\begin{equation}
	f_{throw}(\mathcal{D}_{eval}) = \frac{1}{k}\sum_{\mathcal{D}_{eval}} \Vert r^g_{intersect} - r^g_{intersect}\Vert
\end{equation}
with $r^d_{intersect}, r^g_{intersect}$ being the demonstration and generated throw intersect positions respectively. Seeing as not every model would output a release signal that crossed the threshold (0.5) for every trajectory, this error term was set to be infinite in cases when no throw was defined.

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amsmath}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{cleveref}       % smart cross-referencing
\usepackage{lipsum}         % Can be removed after putting your text content
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{doi}
\usepackage{subcaption}
\usepackage{upgreek}

@@

In this article, an \emph{agent} can be taken to mean the part of the system that acts based on state or observation vectors $\mathbf{s}$ in an \emph{environment}, according to a \emph{policy} $\pi$ which is specified by a parametric \emph{model} -- a function $\pi_{\boldsymbol{\uptheta}}({\mathbf{s}})$ with parameters $\boldsymbol{\uptheta}$ tuned in optimization and produces an output \emph{action} $a$. For all practical intents and purposes, this means that references to the agent, model, and policy are almost interchangeable in most contexts. A formalism common in both reinforcement and imitation learning contexts is the \emph{Markov decision process} (MDP), formally given by \citep{attia2018global}

\begin{equation}
	MDP = (S,A,T,R,I)
\end{equation}
where $S$ is the set of states $s$, $A$ is the (formally discrete) set of actions $a$, $T:S \times A \rightarrow S$ is a state transition function that encapsulates the environment, $R:S \rightarrow \mathbb{R}$ is a reward function associated with each state and $I = p\left(\mathbf{s}_0 \in S\right)$ represents the initial state distribution. In many imitation learning-related cases a reward function need not be specified or considered. Moreover, if one is willing to break with the strict formal definition and give up the use of mathematical tools defined only on finite probability distributions, continuous action spaces can also be considered.

@@

\begin{equation}
	t, k, n \in \mathbb{N}; \pi_{\boldsymbol{\uptheta}}:\lbrace {(\mathbf{s}_n)_{t-k}^t} \rbrace \rightarrow A
\end{equation}
which, by allowing for input sequences of variable length, may become a function defined on the entire known state history:

\begin{equation}
	\pi_{\boldsymbol{\uptheta}}:\lbrace {(\mathbf{s}_n)_1^t} \rbrace \rightarrow A
\end{equation}

@@

\begin{equation}
	\pi_{\boldsymbol{\uptheta}}:\lbrace {(\mathbf{s}_n)_1^t} \rbrace \rightarrow S
\end{equation}
which, when running the model on its prior output, is equivalent to sequence-building tasks encountered in domains such as text generation.

@@

After each recording session, an entirely unstructured corpus is available. It contains timestamped pose observations collected at whatever intervals they happened to have been generated by the tracking software. Each observation only contains information about a single tracked body. A peculiarity of this particular recording method is that observations about each body tend to arrive in bursts:
\begin{equation}
    \mathbf{s}^a_{t1}, \mathbf{s}^a_{t2}, ..., \mathbf{s}^b_{tm-1}, \mathbf{s}^b_{tm}
\end{equation}
where $\mathbf{s}^a_{t \in \mathbb{R}}$ corresponds to the observed state of object $a$ at time $t$.
Thus, to obtain sequential data in discrete time, it is necessary to resample the positions and orientations of all relevant bodies at a constant time step:
\begin{equation}
    (\mathbf{s}^a_{1}, \mathbf{s}^b_{1}, \mathbf{s}^c_{1}), ..., (\mathbf{s}^a_{k}, \mathbf{s}^b_{k}, \mathbf{s}^c_{k}) = \mathbf{s}_1, ..., \mathbf{s}_k
\end{equation}

@@

\begin{equation}
	(\mathbf{s}_{t_{start}}, ... , \mathbf{s}_{t_{start}+steps}), steps \in \mathbb{N}
\end{equation}

@@

\begin{equation}
	(\mathbf{r}_{t+1}, \mathbf{q}_{t+1}, g_{t+1}) = \pi_{\boldsymbol{\uptheta}}\left(\frac{t}{f_{sample}}, \mathbf{r}_{t}, \mathbf{q}_{t}, g_{t}, \mathbf{r}^{target}_t \right)
\end{equation}
where $\mathbf{r}_{t}, \mathbf{q}_{t}, g_{t}$ represent the end effector translation vector, orientation quaternion and gripper actuator signal respectively at time step $t$ in both the input and output. The input is augmented with a time/phase signal (discrete time step divided by the sampling frequency, in this case $100Hz$) relative to the start of the demonstration or generated trajectory, as well as the target coordinate vector $\mathbf{r}^{target}_t$ which corresponds to the extrapolated target coordinates in the demonstration data set and commanded throw coordinates at inference. The time signal was added to the input as it was found that trajectories generated by feedforward networks were liable to diverge without it, and the data sets thus modified were used for all training thereafter. The gripper control signal has values $\lbrace 0,1 \rbrace$ in the training data set.

@@

\begin{equation}
	\mathcal{D}_{eval} = \left ( (\tau^d_{1}, \tau^g_{1}), ..., (\tau^d_{k}, \tau^g_{k}) \right )
\end{equation}
\begin{equation}
	\tau^d, \tau^g = (\mathbf{s}^d_1, ..., \mathbf{s}^d_m), (\mathbf{s}^g_1, ..., \mathbf{s}^g_m); \mathbf{s}^d_1 = \mathbf{s}^g_1 
\end{equation}
where $\mathcal{D}_{eval}$ refers to the evaluation data set of a single model with respect to either the validation or test demonstration set, but $\tau^d_i, \tau^g_i$ are the demonstration and generated trajectories (sequences of state observations $s_j$) sharing the same initial state respectively.

@@

\begin{equation}
	f_{stepwise}(\mathcal{D}_{eval}) = f: \left [ \sum_{\tau^d, \tau^g \in \mathcal{D}_{eval}}\sum_{i=1}^m \left ( g:\mathbf{s}^d_i \times \mathbf{s}^g_i \rightarrow \mathbb{R} \right ) \right ] \rightarrow \mathbb{R}
\end{equation}
where the inner function $g$ corresponds to metrics such as position error, rotation error (quaternion angular distance), or categorical cross-entropy in the release signal. Finally, a release error metric was computed trajectory-wise. To do this, the release point of each trajectory was found, the corresponding position found and the velocity vector estimated. When considering ways to combine these two terms into a single quantitative error estimator, it was decided that a simple ballistic extrapolation and resulting miss distance along the ground plane would serve as a decent first-order approximation of throw accuracy \emph{vis-à-vis} the training or validation data set.

@@

\begin{equation}
	\mathbf{r}_{intersect} = \left ( x(t_{intersect}), y(t_{intersect}), z(t_{intersect}) \right )
\end{equation}
and the trajectory-wise throw error metric computed by:

\begin{equation}
	f_{throw}(\mathcal{D}_{eval}) = \frac{1}{k}\sum_{\mathcal{D}_{eval}} \Vert \mathbf{r}^g_{intersect} - \mathbf{r}^g_{intersect}\Vert
\end{equation}
with $\mathbf{r}^d_{intersect}, \mathbf{r}^g_{intersect}$ being the demonstration and generated throw intersect positions respectively. Seeing as not every model would output a release signal that crossed the threshold (0.5) for every trajectory, this error term was set to be infinite in cases when no throw was defined.
